---
title: "dp_cdf_pipline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dp_cdf_pipline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dpBootstrap)
```

Here is the data that we want to use in the example. It's a sample from a (standard)-lognormal distribution.

```{r}
data <- exp(rnorm(1000))
```


First you need to define the differentially private mechanism that you want to apply to the data.
Note that for this particular dp cdf mechanism you need to provide a lower and upper bound where you expect the data to be (without looking at the data!).

```{r}
mechanism <- function(x) {
  lower_bound <- 0
  upper_bound <- 40
  
  # This generates a projected dp cdf of the data. All functions are part of the
  # dpBootstrap package.
  projected_dp_cdf <-
    x |> bin_data(lower_bound = lower_bound, upper_bound = upper_bound) |> dp_cdf(rho = 0.5) |>
    project_dp_cdf()
  
  # For the generic pipeline to work we want the output of the mechanism to be
  # synthetic data in the same format as the input data.
  M_inv <- get_inv_M(length(projected_dp_cdf$y))
  
  # Get counts per bin
  tmp_count <-
    round(M_inv %*% projected_dp_cdf$y * projected_dp_cdf$n, 0)
  
  # Replicate each bin the number of counts (this will bias mean estimation on 
  # the synthetic data by approx. bin_width/2)
  synth_data <- NULL
  for (i in 1:nrow(tmp_count)) {
    synth_data <-
      c(synth_data, rep(projected_dp_cdf$x[i], tmp_count[i,]))
  }
  
  # Make sure that examples in the excess bins are outside the boundaries but
  # not -Inf or Inf (R won't count them otherwise).
  synth_data[synth_data == -Inf] <- lower_bound - 1
  synth_data[synth_data == Inf] <- upper_bound + 1
  
  return(synth_data)
}
```

What statistics of interest do you want to calculate on the private data?
The output should be a named list.

```{r}
statistics_of_interest <- function(x) {
  list(median = median(x),
       mean = mean(x),
       quantiles = quantile(x, c(0.1, 0.25, 0.75, 0.9)))
}
```

How do we get one bootstrap sample from the synthetic data?
Here we use poisson sampling.

```{r}
one_bootstrap_sample <- function(x_tilde) {
  pois <- rpois(length(x_tilde), 1)
  boot_samp <- NULL
  for (i in 1:length(pois)) {
    boot_samp <- c(boot_samp, rep(i, pois[i]))
  }
  return(boot_samp)
}
```

Now we repeat this to generate a list of ids for each bootstrap repetition

```{r}
bootstrap_sampler <- function(x_tilde, B = 1000) {
  boot_ids_list <- lapply(1:B, function(x)
    one_boot_sample(x_tilde))
  return(boot_ids_list)
}
```

And we can resample data and apply the mechanism.

```{r}
bootstrap_loop <- function(x_tilde, bootstrap_ids_list, mechanism, statistics_of_interest) {
  b <- 1
  theta_tilde_tilde_list <- vector("list", length = length(bootstrap_ids_list))
  
  cli::cli_progress_bar("Bootstrapping", total = length(bootstrap_ids_list))
  for(bootstrap_ids in bootstrap_ids_list){
    x_tilde_hat <- x_tilde[bootstrap_ids]
    x_tilde_tilde <- mechanism(x_tilde_hat)
    theta_tilde_tilde <- statistics_of_interest(x_tilde_tilde)
    theta_tilde_tilde_list[[b]] <- theta_tilde_tilde
    b <- b + 1
    cli::cli_progress_update()
  }
  
  return(theta_tilde_tilde_list)
}
```

Now we have everything defined to run the dp_bootstrap function. Note that depending on the input data and mechanism all of the above functions need to be adapted to the application.

```{r}
results <- dp_bootstrap(
  x = data,
  mechanism = mechanism,
  statistics_of_interest = statistics_of_interest,
  B = 1000,
  one_bootstrap_sample,
  bootstrap_sampler,
  bootstrap_loop
)
```

