---
title: "multivariate_synthetic_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multivariate_synthetic_data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dpBootstrap)
library(reticulate)

use_condaenv("r-reticulate")
```


```{r}

# You need to install the private-pgm python package in the conda environment 
# to get started.
mbi <- reticulate::import("mbi")

sp <- reticulate::import("scipy")

full_data <-
  mbi$Dataset$load(
    '/Users/marcelneunhoeffer/Desktop/github/private-pgm/data/adult.csv',
    '/Users/marcelneunhoeffer/Desktop/github/private-pgm/data/adult-domain.json'
  )


data <-
  mbi$Dataset$load(
    '/Users/marcelneunhoeffer/Desktop/github/private-pgm/data/adult.csv',
    '/Users/marcelneunhoeffer/Desktop/github/private-pgm/data/adult-domain.json'
  )


test_data <-
  mbi$Dataset$load(
    '/Users/marcelneunhoeffer/Desktop/github/private-pgm/data/adult.csv',
    '/Users/marcelneunhoeffer/Desktop/github/private-pgm/data/adult-domain.json'
  )


# Take a sample from the data

sel <- sample(nrow(data$df), size = 5000)
data$df <- data$df[sel, ]

testset <- test_data$df

testset <- testset[-sel,]
# Define a draw from a Laplace distribution

rlaplace <- function(n, b) {
  rexp(n, 1 / b) - rexp(n, 1 / b)
}
```

Now we can define the mechanism. It works on the mbi Dataset class and returns an object of the same class.
First the unsupported case (only one-way marginals are measured).

```{r}
mechanism <- function(data) {
  rdomain <- colnames(data$df)
  domain <- data$domain
  total <- dim(data$df)[1]
  
  cliques <- list()
  
  
  # cliques = list(
  #   list('age', 'education-num'),
  #   list('marital-status', 'race'),
  #   list('sex', 'hours-per-week'),
  #   list('hours-per-week', 'income>50K'),
  #   list('education-num', 'income>50K'),
  #   list('native-country', 'marital-status', 'occupation')
  # )
  
  # cliques = list(
  #   list('age', 'education-num'),
  #   list('education-num', 'race'),
  #   list('race', 'hours-per-week'),
  #   list('hours-per-week', 'income>50K'),
  #   list('native-country', 'income>50K'),
  #   list('native-country', 'marital-status', 'occupation')
  # )
  
  
  
  epsilon = 1.0
  epsilon_split = epsilon / (length(domain) + length(cliques))
  sigma = 2.0 / epsilon_split
  
  
  measurements <- list()
  counter <- 1
  for (col in rdomain) {
    x <- data$project(col)$datavector()
    y = x + rlaplace(n = length(x), b = sigma)
    I = Matrix::Diagonal(length(x))
    measurements[[counter]] <- list(I, y, sigma, list(col))
    counter <- counter + 1
  }
  
  
  
  if(length(cliques > 0)){
  for (cl in cliques) {
    x = data$project(cl)$datavector()
    y = x + rlaplace(n = length(x), b = sigma)
    I = Matrix::Diagonal(length(x))
    measurements[[counter]] <- list(I, y, sigma, cl)
    counter <- counter + 1
  }
  }
  engine = mbi$FactoredInference(domain, log=TRUE, iters=200L)
  #engine =  mbi$LocalInference(domain,
  #                             log = TRUE,
  #                             iters = 2500L,
  #                            marginal_oracle = 'convex')
  
  model = engine$estimate(measurements, total = total)
  
  synth = model$synthetic_data(rows = as.integer(total))
  
  return(synth)
  
}
```

What statistics of interest do you want to calculate? Make sure that the calculation works on the output object.
Here just some column statistics (mean and quantiles) and a simple logistic regression.

```{r}
statistics_of_interest <- function(x) {
  
  
  df <- x$df
  
  m1 <- glm(
    `income>50K` ~ age + workclass + `education-num` + `hours-per-week`,
    family = binomial(link = "logit"), data = df
  )
  auc <- as.numeric(pROC::auc(pROC::roc(testset$`income>50K`, predict.glm(m1, newdata = testset, type = "response"), direction = "<")))

  
  list(
    mean = apply(x$df, 2, mean),
    quantiles = apply(x$df, 2, quantile, c(0.1, 0.25, 0.5, 0.75, 0.9)),
    logit_coefs = coef(
      m1
    ),
    auc = auc
  )
}

``` 

Now we can define the bootstrap loop.

```{r}

bootstrap_loop <-
  function(x_tilde,
           mechanism,
           statistics_of_interest,
           B) {
    theta_tilde_tilde_list <- vector("list", length = B)
    
    cli::cli_progress_bar("Bootstrapping", total = B)
    for (b in 1:B) {
      pois <- rpois(nrow(x_tilde$df), 1)
      boot_samp <- NULL
      for (i in 1:length(pois)) {
        boot_samp <- c(boot_samp, rep(i, pois[i]))
      }
      x_tilde_hat <- x_tilde
      x_tilde_hat$df <- x_tilde$df[boot_samp,]
      x_tilde_tilde <- mechanism(x_tilde_hat)
      theta_tilde_tilde <- statistics_of_interest(x_tilde_tilde)
      theta_tilde_tilde_list[[b]] <- theta_tilde_tilde
      cli::cli_progress_update()
    }
    
    return(theta_tilde_tilde_list)
  }
```


```{r}

results_pgm1 <- dp_bootstrap(
  x = data,
  mechanism = mechanism,
  statistics_of_interest = statistics_of_interest,
  B = 10,
  bootstrap_loop
)
```

Finally, summarize the results.

```{r}
# Statistic of interest
results_pgm1$theta_tilde$logit_coefs

results_pgm1$theta_tilde$auc
sapply(results1$theta_tilde_tilde, function(x) x$auc)

# And confidence intervals
apply(sapply(results_pgm1$theta_tilde_tilde, function(x) x$logit_coefs), 1, quantile, c(0.025, 0.975))

statistics_of_interest(data)$logit_coefs
statistics_of_interest(full_data)$logit_coefs


```



```{r}
mechanism <- function(data) {
  rdomain <- colnames(data$df)
  domain <- data$domain
  total <- dim(data$df)[1]
  
  # cliques = list(
  #   list('age', 'income>50K'),
  #   list('workclass', 'income>50K'),
  #   list('education-num', 'income>50K'),
  #   list('hours-per-week', 'income>50K')
  #   #list('education-num', 'income>50K'),
  #   #list('native-country', 'marital-status', 'occupation')
  # )
  
  cliques <- list(
    list("workclass", "education-num",
                 "marital-status", "occupation", "relationship",
                 "race", "sex",
                 "hours-per-week", "income>50K")
  )
  
  # cliques = list(
  #   list('age', 'education-num'),
  #   list('education-num', 'race'),
  #   list('race', 'hours-per-week'),
  #   list('hours-per-week', 'income>50K'),
  #   list('native-country', 'income>50K'),
  #   list('native-country', 'marital-status', 'occupation')
  # )
  
  
  
  epsilon = 1.0
  epsilon_split = epsilon / (length(domain) + length(cliques))
  sigma = 2.0 / epsilon_split
  
  
  measurements <- list()
  counter <- 1
  for (col in rdomain) {
    x <- data$project(col)$datavector()
    y = x + rlaplace(n = length(x), b = sigma)
    I = Matrix::Diagonal(length(x))
    measurements[[counter]] <- list(I, y, sigma, list(col))
    counter <- counter + 1
  }
  
  cl <- cliques[[1]]
  
  for (cl in cliques) {
    x = data$project(cl)$datavector()
    y = x + rlaplace(n = length(x), b = sigma)
    I = Matrix::Diagonal(length(x))
    measurements[[counter]] <- list(I, y, sigma, cl)
    counter <- counter + 1
  }
  
  engine = mbi$FactoredInference(domain, log=TRUE, iters=200L)
  #engine =  mbi$LocalInference(domain,
  #                             log = TRUE,
  #                             iters = 2500L,
  #                            marginal_oracle = 'convex')
  
  model = engine$estimate(measurements, total = total)
  
  synth = model$synthetic_data(rows = as.integer(total))
  
  return(synth)
  
}
```

```{r}

results_pgm2 <- dp_bootstrap(
  x = data,
  mechanism = mechanism,
  statistics_of_interest = statistics_of_interest,
  B = 10,
  bootstrap_loop
)
```

Finally, summarize the results.

```{r}
# Statistic of interest
results_pgm2$theta_tilde$logit_coefs

results_pgm2$theta_tilde$auc
sapply(results_pgm2$theta_tilde_tilde, function(x) x$auc)

# And confidence intervals
apply(sapply(results_pgm2$theta_tilde_tilde, function(x) x$logit_coefs), 1, quantile, c(0.025, 0.975))

statistics_of_interest(data)
statistics_of_interest(full_data)


```
